---
title: "Statswars and Matrices"
description: |
  Relational analysis that says more than 25 000 emotional words
author:
  - name: recraft@pm.me
    url: https://recraft.me
    affiliation: Recraft
    affiliation_url: https://github.com/recraft
date: "2019-02-10"
output:
  radix::radix_article:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = "100%")

```

In a short but sweet article that appeared on [r-bloggers](https://www.r-bloggers.com/statswars/), Kieran Healy uses one image to convey 25 000 emotional words about the asymmetric relationships between 5 different user communities in the data science world -  R, Python, STATA, SPSS and SAS communities. A full resolution version of this single compact ["emo matrix" is published at Kieran Healy's blog](https://kieranhealy.org/blog/archives/2019/02/07/statswars/).

```{r preview=TRUE}
library(magick)

image_read("https://kieranhealy.org/files/misc/package_wars.png") %>% 
image_convert("png") %>% image_convert(colorspace = "gray") %>%
image_resize("x400")
```

How can we create a tabular matrix-like visual like this, filled with images that summarize how a certain community views itself and other communities while also including the opposite perspective - how those other communities views that community from the outside?

## Deconstructing the emo matrix

The data displayed in the emomatrix is "relational data" - data describing a relationship between sets of two communities. For instance: the python community sees the R community in certain way that evokes emotions such as those inspired by a particular image.

We start deconstructing the emo matrix but breaking out the sub images so we can learn more about it.

```{r, echo=TRUE}

library(magick)
library(purrr)

emomatrix <- 
  "https://kieranhealy.org/files/misc/package_wars.png" %>%
  image_read()

imagemagick_geometries <- function() {
  # assemble ImageMagick geometry strings
  # data for image regions to crop
  # :) inspired by TI Extended Basic DATA syntax
  # http://www.99er.net/files/TI%20Extended%20Basic%20-%20Linked.pdf 
  widths <- c(400, 400, 300, 278, 278)
  heights <- c(270, 270, 270, 270, 240)
  offsets_w <- cumsum(c(266, 400, 440, 350, 300))
  offsets_h <- cumsum(c(226, 270, 270, 270, 270))
  paste_plus <- function(x, y) paste0("+", paste(x, y, sep = "+"))
  paste_x <- function(x, y) paste0(x, "x", y)
  xy <- outer(offsets_w, offsets_h, paste_plus)
  wh <- outer(widths, heights, paste_x)
  whxy <- paste0(wh, xy)
  dim(whxy) <- c(5, 5)
  return (whxy)
}

# sub-images
five <- c("STATA", "R", "SAS", "Python", "SPSS")
geoms <- imagemagick_geometries()
emoimages <- map(geoms, function(x) image_crop(emomatrix, x))
dim(emoimages) <- c(5, 5)
colnames(emoimages) <- rownames(emoimages) <- five
```

We can now inspect a particular relation from both ends. 

For example, how does the R community view the Python community. The image seems to indicate a mostly positive view, with some envy in it?

```{r, echo=TRUE}
emoimages["R", "Python"][[1]]
```

From the other end we can see how the Python community views the R community. If you're Homer - wouldn't you like to have and use Homer's Dream Car - it has all the features you would ever need? There is some ambiguity however - or don't you perceive a hint of ridicule being present there?

```{r, echo=TRUE}
emoimages["Python", "R"][[1]]
```

Each image says a thousand emotional words in the form of inside jokes - not easy to decode for outsiders to the communities involved, but with a lot of emotional content that can be quickly be deciphered by members inside these communities. 

For an outsider, at this point, we'd have to use the individual images for example with a reverse image search API to locate their origins in order to learn more about a specific view and to try to get to the gist of what a particular image is trying to communicate.

```{r, eval=FALSE, echo=TRUE}
res <- map(whxy, function(x) image_crop(emomatrix, x) %>% 
  image_resize("x240") %>% image_write(paste0("statswars-", x, ".png")))
```


## The emotional matrix reconstructed

Here follows an attempt to reproduce this diagram using `DT` which allows us to get a HTML table that can link to images and provide tooltips that add some clarifications.

```{r echo=TRUE}

library(DT)
library(rvest)
library(httr)
library(tibble)
library(stringr)
library(dplyr)
library(readr)
library(knitr)
library(kableExtra)

imagedata <- read_delim(delim = ";", trim_ws = TRUE, file = "emo; url
IBM Business Automation Content Analyzer on Cloud; https://www.bpm.ibmcloud.com/images/baca/card3.jpg
A Rube Goldberg Machine; https://upload.wikimedia.org/wikipedia/commons/a/a9/Rube_Goldberg%27s_%22Self-Operating_Napkin%22_%28cropped%29.gif
IBM 1800 DACS; http://www.opendragon.ca/Images/001_05_12_0006.jpg
Tech Bro Says You Just Do Not Understand the Blockchain; https://mondrian.mashable.com/uploads%252Fcard%252Fimage%252F877740%252F87e3e914-63e0-43f1-a53a-5d4cb1b12cc8.jpg%252F950x534__filters%253Aquality%252890%2529.jpg?signature=vin5OtSuozuc6naKbxCh0A85xhU=&source=https%3A%2F%2Fblueprint-api-production.s3.amazonaws.com
YMCA; https://www.ymcala.org/sites/default/files/styles/node_teaser/public/2017-11/Graduates.jpg?itok=S2kLygK0
Giphy Monkey Stuff; http://cdn.lowgif.com/full/ba25ae8ed81cebe8-monkey-typing-gifs-tenor.gif
The Matrix 4 Rebooted; https://fsmedia.imgix.net/25/59/a1/34/315d/495a/9c0c/66b36664c23c/the-matrix6jpg.jpeg?rect=0%2C0%2C1920%2C960&auto=format%2Ccompress&w=650
SAGE; https://image.slidesharecdn.com/startsmall-staysmall-130829072518-phpapp01/95/start-small-stay-small-4-638.jpg?cb=1377768482
Distracted Boyfriend; https://proxy.duckduckgo.com/iu/?u=https%3A%2F%2Fi.imgflip.com%2F1v1xf4.jpg&f=1
Spongebob and Patrick Are Best Friends Forever; https://static.artfire.com/uploads/images/561/45f/d3b/bee/large/56145fd3bbeea01212bff0e3a516eb506cc7ad63335596adb76006e5bcfd6ed0.jpg
European Teenage Business Man; https://image.shutterstock.com/image-photo/european-appearance-teenager-boy-business-450w-326683721.jpg
Avalon - Where Excalibur Gets Forged - Messy Computer Desk; https://c7.alamy.com/comp/A3NY2A/messy-computer-desk-A3NY2A.jpg
Analytics Dashboard; https://www.rythmos.com/hs-fs/hubfs/BI%20TOOLS%20BLOG.jpg?width=600&name=BI%20TOOLS%20BLOG.jpg
Tech Bro Says You Just Do Not Understand the Blockchain; https://mondrian.mashable.com/uploads%252Fcard%252Fimage%252F877740%252F87e3e914-63e0-43f1-a53a-5d4cb1b12cc8.jpg%252F950x534__filters%253Aquality%252890%2529.jpg?signature=vin5OtSuozuc6naKbxCh0A85xhU=&source=https%3A%2F%2Fblueprint-api-production.s3.amazonaws.com
Baby with Laptop Feels Confused; http://image.shutterstock.com/display_pic_with_logo/237994/237994,1315555947,1/stock-photo-baby-with-laptop-confused-84320152.jpg
I Hate New Excel; https://www.insightsquared.com/wp-content/uploads/2012/04/i-hate-excel.png
The Homer - Your Dream Car With All Features; https://laughingsquid.com/wp-content/uploads/2014/05/14104880051_92d6e1f922_z.jpg
Comptometer; https://upload.wikimedia.org/wikipedia/commons/f/ff/Mechanical-Calculator.png
Tech Bro Says You Just Do Not Understand the Blockchain; https://mondrian.mashable.com/uploads%252Fcard%252Fimage%252F877740%252F87e3e914-63e0-43f1-a53a-5d4cb1b12cc8.jpg%252F950x534__filters%253Aquality%252890%2529.jpg?signature=vin5OtSuozuc6naKbxCh0A85xhU=&source=https%3A%2F%2Fblueprint-api-production.s3.amazonaws.com
Correct Execution of the Process; https://image.shutterstock.com/image-vector/operator-controlling-correct-execution-process-450w-186044081.jpg
I Hate Old Excel; https://www.insightsquared.com/wp-content/uploads/2012/04/i-hate-excel.png
IBM Selective Sequence Electronic Calculator (SSEC)- Electromechanical Computing technology from 1959; https://slideplayer.fr/slide/3696889/12/images/13/1959.jpg
HAL 900 Eye; http://www.synthstuff.com/mt/archives/2008/Eye_HAL9000.jpg
African Rock Python; https://upload.wikimedia.org/wikipedia/commons/8/80/Python_sebae_head2a.jpg
Homework for Schoolkids; https://www.quill.com/is/image/Quill/JV53677_s7?$img400$
")

# fcn to generate base64 encoded local image linked to external source
img_uri <- function(url, title) { 
  
  x <- tempfile(fileext = ".png")
  
  path <- image_read(url) %>% 
    image_convert(format = "png") %>% 
    image_resize("64x64^") %>% 
    image_write(x)
  message("writing image at ", x, " got return ", path)
  
  sprintf('<a href="%s"><img height=64 title="%s" src="%s"/></a>', 
          url, title, knitr::image_uri(x)) 
}


# add the from, to relation as well as tooltip data and a link to the image url
id <- imagedata %>% bind_cols(data_frame(
    alt = as.vector(outer(five, five, function(x, y) paste0(x, " as seen by ", y))),
    from = as.vector(outer(five, five, function(x, y) paste0(x))),
    to = as.vector(outer(five, five, function(x, y) paste0(y)))
  )) #%>%
#  mutate(img = img_uri(url, emo))
#  mutate(imgurl = paste0("<a href='", url, "'><img height=52 title='", emo, "' src='", url, #"'/></a>"))

img <- purrr::map2(id$url, id$emo, img_uri)
id$img <- unlist(img)

id %>% 
  select(alt, from, to, emo) %>%
  filter(from == "R", to == "R") %>% 
  mutate(url = "https://some.imageurl.here/image.png") %>%
  kable() %>% 
  kable_styling()

```

With the source data expressed like above - with relationships added (from, to) in addition to the image urls - we can use `tidyr` to spread this tall table of data into a format that is suitable for being visualized with `DT` as a webfriendly table of image thumbnails, where hovering will provide extra info through a tooltip and where clicking leads to the source image.

```{r}

library(tidyr)

res <- id %>% 
  select(from, to, img) %>% 
  spread(from, img) %>% 
  rename(Community = to)

datatable(res, escape = FALSE, rownames = FALSE, filter = 'none', class = "compact",
          options = list(searching = FALSE, paging = FALSE, ordering = FALSE, info = FALSE,
                         columnDefs = list(list(className = 'dt-left', targets = 0:5))))
```

## The meme matrix

These days memes and emojis are increasingly used as the universal esperanto of cross-cultural communication so here is an attempt at a meme matrix involving the R and Python communities.

```{r, echo=TRUE}
library(gt)

memedata <- read_delim(delim = ";", trim_ws = TRUE, file = "from;to;imgurl
R;R;https://memegen.link/morpheus/once_you_enter_the_tidyverse_there_is_no_going_back/hadley,_yihui_and_cheng_took_the_red_pill_too.jpg
Python;R;https://memegen.link/keanu/what_if_purrr,_furrr_and_dplyr/destructures_the_mind~q.jpg
Python;Python;https://memegen.link/hipster/tech_bro_understands_the_internets/is_pythonista.jpg
R;Python;https://memegen.link/scc/the_good_python_ds_things_--_pandas,_feather_and_ursa_labs/hadley~q.jpg")

res <- 
  memedata %>% 
  mutate(alt = paste0(to, " as seen by ", from)) %>%
  mutate(html = paste0("<a href='", imgurl, "'><img height=200 title='", alt, "' src='", imgurl, "'/></a>")) %>%
  select(from, to, html) %>%
  spread(from, html) %>% 
  rename(Community = to)


datatable(res, escape = FALSE, rownames = FALSE, filter = 'none', class = "compact",
          options = list(searching = FALSE, paging = FALSE, ordering = FALSE, info = FALSE,
                         columnDefs = list(list(className = 'dt-left', targets = 0:2))))
```

SPSS, SAS and STATA communities are intentionally left out as these communites represent users of specific commercial applications rather than users of open source data science languages :) 

  